---
title: "Part II: Analysis and Forecast Modelling"
author: Sally Chen
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
header-includes:
  \usepackage{fvextra}
  
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, message = FALSE)

# rmarkdown::render('RMarkdown_code.Rmd', output_file = 'RMarkdown_report.pdf', envir = new.env())
```


```{r}
library(fpp)
library(fpp2)
library(forecast)
library(GGally)
library(knitr)
```


```{r}
# Load cleaned data
full = read.csv("full.csv")
train = read.csv("train.csv")
test = read.csv("test.csv")
```


# Defining Weather Elements as Time Series Objects
```{r}
ts_full = ts(full, frequency = 365.25, start = c(1999,230))
ts_train = ts(train, frequency = 365.25, start = c(1999,230))
ts_test = ts(test, frequency = 365.25, start = c(2016,55))

ts_max = ts_train[,"max"]
ts_min = ts_train[,"min"]
ts_sunshine = ts_train[,"sunshine"]
ts_wind = ts_train[,"wind"]
ts_rainfall = ts_train[,"rainfall"]

```


# Train and Test Data Exploration
```{r, message=FALSE}
# Check variables correlations
GGally::ggpairs(train[,1:5])


## Time series plots for weather elements
elements = colnames(ts_train)
par(mfrow=c(2,3))
for (i in 1:5){
  ts.plot(ts_train[,i], type = "l", ylab = elements[i])
}

```

## Inspect seasonal patterns from weather elements
```{r}
# Sunshine -> seasonality, no trend
fitstl_sunshine = stl(ts_sunshine, t.window = 365.25, s.window = 365.25)
autoplot(fitstl_sunshine)
par(mfrow=c(1,2))
acf(remainder(fitstl_sunshine), main = "ACF of residual from seasonal decomposition")
pacf(remainder(fitstl_sunshine), main = "PACF of residual from seasonal decomposition")

# wind -> seasonality, no trend
fitstl_wind = stl(ts_wind, t.window = 365.25, s.window = 365.25)
autoplot(fitstl_wind)
par(mfrow=c(1,2))
acf(remainder(fitstl_wind), main = "ACF of residual from seasonal decomposition")
pacf(remainder(fitstl_wind), main = "PACF of residual from seasonal decomposition")

# rainfall -> no seasonality, no trend
fitstl_rainfall = stl(ts_rainfall, t.window = 365.25, s.window = 365.25)
autoplot(fitstl_rainfall)
par(mfrow=c(1,2))
acf(remainder(fitstl_rainfall), main = "ACF of residual from seasonal decomposition")
pacf(remainder(fitstl_wind), main = "PACF of residual from seasonal decomposition")
```


## Analyse trend in daily minimum and maximum temperature
```{r}
## use moving average
par(mfrow=c(2,1))
plot(ma(ts_max, order = 365), main = "Moving Average for Max Temperature", ylab = expression(degree~C))
plot(ma(ts_min, order = 365), main = "Moving Average for Min Temperature", ylab = expression(degree~C))
```

# Building models to predict maximum temperature
## 1. Linear model with a Fourier term to capture seasonality
```{r}
# Fit a linear trend with seasonality using tslm()
tslm_max = tslm(ts_max ~ trend+ts_rainfall+ts_sunshine+ts_wind+fourier(ts_max,K=2))
summary(tslm_max)
AIC(tslm_max) # AIC = 33683.01

plot(ts_max, col = "grey", main = "Fitted Value from Linear Model", ylab = expression(degree~C))
lines(tslm_max$fitted.values)

## Model diagnosis
par(mfrow=c(2,1))
acf(tslm_max$residuals, main = "ACF of residual from fitted model") # lag q = 2
pacf(tslm_max$residuals, main = "PACF of residual from fitted model") # lag p = 2

Box.test(tslm_max$residuals, type = "Lj") #autocorrelation different from 0
dwtest(tslm_max, alternative = "two")
bgtest(tslm_max, 20)

```

## 2. Dynamic Regression with ARIMA(2,0,2) error
```{r}
# Fit a dynamic regression to capture the dynamics left in the residuals
dr_max = Arima(ts_max, xreg = cbind(ts_rainfall,ts_sunshine,ts_wind,fourier(ts_max,K=2)), order=c(2,0,2))
summary(dr_max) # AIC = 32213.94

plot(ts_max, col = "grey", main = "Fitted Value from Dynamic Regression with ARIMA(2,0,2) errors", ylab = expression(degree~C))
lines(dr_max$fitted)

par(mfrow=c(2,1))
acf(dr_max$residuals, main = "ACF of residual from fitted model")
pacf(dr_max$residuals, main = "PACF of residual from fitted model")
Box.test(dr_max$residuals)

## Construct a function to calculate p-value for fitted models
p_value = function(model){
  t_fit = model$coef/(sqrt(diag(model$var.coef)))
  p_fit = 2*pnorm(abs(t_fit), mean = 0, sd = 1, lower.tail = FALSE)
  
  return(p_fit)
}
p_value(dr_max)


# Forecast using fitted model
fcast_rainfall = forecast(ts_rainfall, method = "ets", h=1508)
fcast_sunshine = forecast(ts_sunshine, method="ets", h=1508)
fcast_wind = forecast(ts_wind, method="ets", h=1508)

fcast_xreg = cbind(fcast_rainfall$mean, fcast_sunshine$mean, fcast_wind$mean, fourier(ts_max,K=2,h=1508))
colnames(fcast_xreg) = names(dr_max$coef)[-c(1:5)]

fcast1 = forecast(dr_max, xreg = fcast_xreg, h=1508)
par(mfrow=c(1,1))
plot(fcast1)
accuracy(fcast1, ts_test[,"max"])
```

## 3. Dynamic Regression with ARIMA(5,1,0) error
```{r}
# Fit another dynamic regression using auto arima
dr2_max = auto.arima(ts_max, xreg = cbind(ts_rainfall,ts_sunshine,ts_wind,fourier(ts_max,K=2)))

summary(dr2_max) # ARIMA(5,1,0), AIC = 32894.15

plot(ts_max, col = "grey", main = "Fitted Value from Dynamic Regression with ARIMA(5,1,0) errors", ylab = expression(degree~C))
lines(dr2_max$fitted)

par(mfrow = c(2,1))
acf(dr2_max$residuals, main = "ACF of residual from fitted model")
pacf(dr2_max$residuals, main = "PACF of residual from fitted model")
Box.test(dr2_max$residuals)


# Forecast using fitted model
fcast2 = forecast(dr2_max, xreg = fcast_xreg, h=1508)
par(mfrow=c(1,1))
plot(fcast2)
accuracy(fcast2, ts_test[,"max"])
```

## 4. Dynamic Regression with ARIMA(2,0,2) error and lag predictors
```{r}
# Fit a dynamic regression with lag values
len_train = length(ts_train[,"max"])
lag1_sunshine = c(NA,ts_sunshine[1:(len_train-1)])
lag2_sunshine = c(rep(NA,2),ts_sunshine[1:(len_train-2)])
lag1_rainfall = c(NA,ts_rainfall[1:(len_train-1)])
lag2_rainfall = c(rep(NA,2),ts_rainfall[1:(len_train-2)])
lag1_wind = c(NA,ts_wind[1:(len_train-1)])
lag2_wind = c(rep(NA,2),ts_wind[1:(len_train-2)])

drlag_max = Arima(ts_max, xreg = cbind(ts_rainfall,lag1_rainfall,lag2_rainfall, ts_sunshine,lag1_sunshine,lag2_sunshine,ts_wind,lag1_wind,lag2_wind, fourier(ts_max,K=2)), order=c(2,0,2))
summary(drlag_max) #AIC = 31560.4

plot(ts_max, col = "grey", main = "Fitted value from Dynamic Regression with ARIMA(2,0,2) errors", ylab = expression(degree~C))
lines(drlag_max$fitted)

par(mfrow=c(2,1))
acf(drlag_max$residuals[3:len_train], main = "ACF of residual from fitted model")
pacf(drlag_max$residuals[3:len_train], main = "PACF of residual from fitted model")
Box.test(drlag_max$residuals)


# Forecast using fitted model
fcast_lag1_rainfall = c(ts_sunshine[6034],fcast_rainfall$mean[1:1507])
fcast_lag2_rainfall = c(ts_sunshine[6033:6034],fcast_rainfall$mean[1:1506])
fcast_lag1_sunshine = c(ts_sunshine[6034],fcast_sunshine$mean[1:1507])
fcast_lag2_sunshine = c(ts_sunshine[6033:6034],fcast_sunshine$mean[1:1506])
fcast_lag1_wind = c(ts_wind[6034],fcast_wind$mean[1:1507])
fcast_lag2_wind = c(ts_wind[6033:6034],fcast_wind$mean[1:1506])

fcast_lag_xreg = cbind(fcast_rainfall$mean,fcast_lag1_rainfall,fcast_lag2_rainfall,fcast_sunshine$mean,fcast_lag1_sunshine,fcast_lag2_sunshine,fcast_wind$mean,fcast_lag1_wind,fcast_lag2_wind, fourier(ts_max,K=2,h=1508))
colnames(fcast_lag_xreg) = names(drlag_max$coef)[-c(1:5)]

fcast3 = forecast(drlag_max, xreg = fcast_lag_xreg, h=1508)
par(mfrow=c(1,1))
plot(fcast3)
accuracy(fcast3, ts_test[,"max"])
```


## Use the best model (No.4) to build full temperature forecast
```{r}
full_len = length(ts_full[,"max"])
full_max = ts_full[,"max"]
full_sunshine = ts_full[,"sunshine"]
full_rainfall = ts_full[,"rainfall"]
full_wind = ts_full[,"wind"]

full_lag1_sunshine = c(NA,full_sunshine[1:(full_len-1)])
full_lag2_sunshine = c(rep(NA,2),full_sunshine[1:(full_len-2)])
full_lag1_rainfall = c(NA,full_rainfall[1:(full_len-1)])
full_lag2_rainfall = c(rep(NA,2),full_rainfall[1:(full_len-2)])
full_lag1_wind = c(NA,full_wind[1:(full_len-1)])
full_lag2_wind = c(rep(NA,2),full_wind[1:(full_len-2)])

finaldr_max = Arima(full_max, xreg = cbind(full_sunshine,full_lag1_sunshine,full_lag2_sunshine, full_rainfall,full_lag1_rainfall,full_lag2_rainfall, full_wind,full_lag1_wind,full_lag2_wind, fourier(full_max,K=2)), order = c(2,0,2))
summary(finaldr_max)

plot(full_max, col = "grey", main = "Fitted value from predictive model for max temperature", ylab = expression(degree~C))
lines(finaldr_max$fitted)

par(mfrow=c(2,1))
acf(finaldr_max$residuals[3:full_len], main = "ACF of residual from fitted model")
pacf(finaldr_max$residuals[3:full_len], main = "PACF of residual from fitted model")
Box.test(finaldr_max$residuals)


## build forecast
fcast_full_sunshine = forecast(full_sunshine, method = "ets", h = 365)
fcast_full_rainfall = forecast(full_rainfall, method = "ets", h = 365)
fcast_full_wind = forecast(full_wind, method = "ets", h = 365)
fcast_full_lag1_sunshine = c(full_sunshine[7542],fcast_full_sunshine$mean[1:364])
fcast_full_lag2_sunshine = c(full_sunshine[7541:7542],fcast_full_sunshine$mean[1:363])
fcast_full_lag1_rainfall = c(full_rainfall[7542],fcast_full_rainfall$mean[1:364])
fcast_full_lag2_rainfall = c(full_rainfall[7541:7542],fcast_full_rainfall$mean[1:363])
fcast_full_lag1_wind = c(full_wind[7542],fcast_full_wind$mean[1:364])
fcast_full_lag2_wind = c(full_wind[7541:7542],fcast_full_wind$mean[1:363])

fcast_full_xreg = cbind(fcast_full_sunshine$mean,fcast_full_lag1_sunshine,fcast_full_lag2_sunshine, fcast_full_rainfall$mean,fcast_full_lag1_rainfall,fcast_full_lag2_rainfall, fcast_full_wind$mean, fcast_full_lag1_wind, fcast_full_lag2_wind, fourier(full_max,K=2, h=365))
colnames(fcast_full_xreg) = names(finaldr_max$coef)[-c(1:5)]

fcast_full_max = forecast(finaldr_max,xreg = fcast_full_xreg, h=365)
par(mfrow=c(1,1))
plot(fcast_full_max)


max_temperature_result = data.frame(seq(as.Date("2020-04-11"), by="days", length.out = 365), round(fcast_full_max$mean,2))
colnames(max_temperature_result) = c("Date", "Forecasted Max Temperature")



kable(max_temperature_result)

```


# Building models to predict the number of extreme heat days


